A Recipe for Scavenging - the natural history of a behaviour

### 2016-04-22 natural history of scavenging.tex
minor edits to the intro

\affiliations{
\item A. Kane (\url{adam.kane@ucc.ie}), University College Cork, Cooperage Building, School of Biological Earth and Environmental Sciences, Cork, Ireland.
\item K. Healy and A. Jackson, Trinity College Dublin, Department of Zoology, School of Natural Sciences, Dublin Ireland.
\item T. Guillerme, Imperial College London, Silwood Park Campus, Department of Life Sciences, Buckhurst Road, Ascot SL5 7PY, United Kingdom.
\item G. Ruxton, School of Biology, Sir Harold Mitchell Building, Greenside Place, St Andrews, KY16 9TH, United Kingdom.
}

@@

Historically, scavengers have not been viewed as the most charismatic of animals.
This may go some way to explaining the gap in our knowledge of the prevalence of this behaviour \citep{devault2003scavenging}.
Professor Sanborn Tenney writing in 1877 for The American Naturalist had this to say about one well known group, ``prominent among the mammalian scavengers are the hyenas, the ugliest in their general appearance of all the flesh eaters."
He contrasts these with ``nobler kinds" of carnivores such as lions and tigers \citep{tenney1877few}.
Even aside from our own subjective biases, scavenging is a difficult behaviour to detect after the fact.
Without catching a carnivore in the act of killing we are left to infer how the prey was killed.
Some simple heuristics can inform us, for instance, in cases where the prey item was simply too large to have been killed by the ostensible predator \citep{pobiner2008paleoecological}.
But clearly, a scavenger doesn't only feed on animals too big for it to have hunted.
The obvious lack of direct behavioural data compounds the difficulty of discerning scavenging among extinct forms.
Indeed, a single species of dinosaur notwithstanding \citep{carbone2011intra}, a synthesis describing the natural history of scavengers is absent from the literature.
%TG: I don't understand the dinosaur part of this sentence. Might be my non-native english but I thought it's worth noting for other non-english folk out there (even though I know they aren't many!).
Fortunately, research on scavenging is on the rise \citep{manga2006vulture}.
As a result, we are now beginning to realise the extent of this behaviour such that, ``in some ecosystems, vertebrates have been documented to assimilate as much as 90\% of the available carrion" \citep{beasley2015vertebrates}.
This has profound implications for the trophic ecology of these systems and particularly our models of them.
Even Tenney’s noble big cats are now known to take in a significant portion of carrion in their diet where some lion populations get over 50\% of their meat from carcasses \citep{jones2015african}.
By recognising the difficulty in directly observing scavenging, other methods have been turned to to discern the most suitable morphologies, physiologies and environments for a scavenging lifestyle to prosper.
Here we chart the natural history of scavenging by assessing the potential for the behaviour in dominant vertebrate groups given their ecology and functional traits.
%It is our aim in this review to employ these methods to gain an understanding of scavengers past and present.

@@

\subsection*{The Challenges of Scavenging} %TG: challenges or difficulty? up to you.
The chief hurdle to scavenging is finding a resource that is often difficult to predict in space and time.
Through chance alone many species will avail of some opportunistic scavenging. %TG: Two times "opportunities", replaced by "chances"
However, species that rely on scavenging to sustain substantial portions of their diets must increase the probability of encountering a sufficient amount of carrion in order to meet their energetic demands.
%Once found, the scavenger must be able to out-compete any potential competitors and process a product under the processes of decay from toxin producing and disease causing microorganisms \citep{ruxton2014fruit}. %TG: make the following sentence more sense?
Once found, the scavenger must be able to out-compete any potential  competitors and process the, typically decaying, carcass replete with microorganism derived toxins  \citep{ruxton2014fruit}.
Finally, the potential for scavenging will also depend on the density, size, and quality of carcases produced, all of which are affected by complex ecosystem dynamics.
All of these facets are essentially the key parameters found in functional response curves, namely encounter rate, handling time and prey availability \citep{jeschke2002predator}. %TG: Jaysus, this sentence doesn't read well! "Functional response curves of optimal foraging theory" can probably be phrased in a nicer/easier way...
%With this framework we compare the ability of scavenging across vertebrates. % focusing on each of these facets and their relationship with scavenging.
By considering scavenging in this context of optimal foraging we can identify the prerequisite attributes and processes required for the behaviour. 
This has enabled us to propose a scale of scavenging whereupon we can place any vertebrate species, past or present, and assess the importance of carrion in its diet.

>>>>

\affiliations{
\item A. Kane (\url{adam.kane@ucc.ie}), University College Cork, Cooperage Building, School of Biological Earth and Environmental Sciences, Cork, Ireland.
\item K. Healy and A.L. Jackson, Trinity College Dublin, Department of Zoology, School of Natural Sciences, Dublin 2, Ireland; Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland.
\item T. Guillerme, Imperial College London, Silwood Park Campus, Department of Life Sciences, Buckhurst Road, Ascot SL5 7PY, United Kingdom.
\item G. Ruxton, School of Biology, Sir Harold Mitchell Building, Greenside Place, St Andrews, KY16 9TH, United Kingdom.
}

@@

Historically, scavengers have not been viewed as the most charismatic of animals.
This may go some way to explaining the gap in our knowledge of the prevalence of this behaviour \citep{devault2003scavenging}.
Professor Sanborn Tenney writing in 1877 for The American Naturalist had this to say about one well known group, ``prominent among the mammalian scavengers are the hyenas, the ugliest in their general appearance of all the flesh eaters."
He contrasts these with ``nobler kinds" of carnivores such as lions and tigers \citep{tenney1877few}.
Even aside from our own subjective biases, scavenging is a difficult behaviour to detect after the fact.
Without catching a carnivore in the act of killing we are left to infer how the prey was killed.
Some simple heuristics can inform us, for instance, in cases where the prey item was simply too large to have been killed by the ostensible predator \citep{pobiner2008paleoecological}.
But clearly, a scavenger does not only feed on animals too big for it to have hunted.
The obvious lack of direct behavioural data compounds the difficulty of discerning scavenging from predation among extinct forms.
Indeed, a single species of dinosaur notwithstanding \citep{carbone2011intra}, a synthesis describing the natural history of scavengers is absent from the literature.
%TG: I don't understand the dinosaur part of this sentence. Might be my non-native english but I thought it's worth noting for other non-english folk out there (even though I know they aren't many!).
With research on scavenging on the rise \citep{manga2006vulture} we are now beginning to realise the extent of this behaviour such that, ``in some ecosystems, vertebrates have been documented to assimilate as much as 90\% of the available carrion" \citep{beasley2015vertebrates}.
This has profound implications for the trophic ecology of these systems and particularly our models of them.
Even Tenney’s noble big cats are now known to take in a significant portion of carrion in their diet where some lion populations acquire over 50\% of their meat from carcasses \citep{jones2015african}.
While recognising the difficulty in directly observing scavenging, it is possible to turn to other methods in order to discern the most suitable morphologies, physiologies and environments for a scavenging lifestyle to prosper.
Here we chart the natural history of scavenging by assessing the potential for the behaviour in dominant vertebrate groups given their ecology and functional traits.
%It is our aim in this review to employ these methods to gain an understanding of scavengers past and present.

@@

\subsection*{The Challenges of Scavenging} %TG: challenges or difficulty? up to you.
The chief hurdle to scavenging is finding a resource that is often difficult to predict in space and time.
Through chance alone many species will avail of some opportunistic scavenging. %TG: Two times "opportunities", replaced by "chances"
However, species that rely on scavenging to sustain substantial portions of their diets must encounter a sufficient amount of carrion in order to meet their energetic demands.
%Once found, the scavenger must be able to out-compete any potential competitors and process a product under the processes of decay from toxin producing and disease causing microorganisms \citep{ruxton2014fruit}. %TG: make the following sentence more sense?
Once found, the scavenger must be able to out-compete any potential  competitors and process the, typically decaying, carcass replete with microorganism derived toxins  \citep{ruxton2014fruit}. Both these characteristics can be assumed to be under evolutionary selection pressures for traits that increase carrion discovery and monopoly.
Finally, the potential for scavenging will also depend on the density, size, and quality of carcases produced, all of which are affected by complex ecosystem dynamics and are outside the selection pressures on the scavenger.
Each of these facets are essentially the backbone of fundamental ecological theory and are the key parameters defined in functional response curves, namely encounter rate, handling time and prey availability \citep{jeschke2002predator}. %TG: Jaysus, this sentence doesn't read well! "Functional response curves of optimal foraging theory" can probably be phrased in a nicer/easier way...
%With this framework we compare the ability of scavenging across vertebrates. % focusing on each of these facets and their relationship with scavenging.
By considering scavenging in this context of optimal foraging we can identify the prerequisite attributes and processes required for the behaviour. 
This has enabled us to propose a scale of scavenging whereupon we can place any vertebrate species, past or present, and assess the importance of carrion in its diet.
